---
title: "Activity 4 - Day 1"
output: github_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Task 2: Load the necessary packages

```{r}
library(tidyverse)
library(tidymodels)
library(GGally)
```

### Task 3: Load the data and

```{r}
evals <- read.delim("https://www.openintro.org/data/tab-delimited/evals.txt")
```

This is an experimental study.

From the histogram output, the distribution of score is the kind of left skewed.

```{r}
# Age vs beauty average
evals %>% ggplot(aes(x = age,
                   y = bty_avg)) + 
  geom_point(alpha=0.20) +
  geom_smooth(se = FALSE, method = "lm", size = 1) +
  labs(title = "Relationship between age and beauty avg",
       x = "Age of Professor",
       y = "Average beauty rating of professor",
       caption = "Data source: The OpenIntro site") + 
  theme_bw()
```

Age appears to have an influence on average beauty rating. It shows a decline in mean beauty rating as age increases. 

```{r}
# Average beauty vs score
evals %>% ggplot(aes(x = bty_avg,
                   y = score)) + 
  geom_point(alpha=0.20) +
  geom_smooth(se = FALSE, method = "lm", size = 1) +
  labs(title = "Relationship between Average Beauty Rating and Evaluation Scores",
       x = "Average Beauty Rating",
       y = "Evalution scores",
       caption = "Data source: The OpenIntro site") + 
  theme_bw()
```

```{r}
# Fitting the model for Average beauty rating and Scores
m_bty <- lm(score ~ bty_avg, data = evals)
tidy(m_bty)

plot(m_bty)
```

### Simplified equation of the line corresponding to score:

Estimated score = beta0hat + beta1hat(bty_avg)
                = 3.88033795 + 0.06663704(bty_avg)

### Interpretation:

Intercept 3.88 is significantly different from zero.
For professors with a beauty average rating of 0, we expect their mean evaluation score to be 3.88.

First slope 0.066 is significantly different from 0. Every 1-unit increase in beauty average rating scale is associated with 0.066 increase in the average evaluation score.

### Task 4: Pairwise relationships

```{r}
# Pairwise relationships
evals %>% 
  select(starts_with("bty_")) %>% 
  ggpairs()
```


I recommend to include "beauty average" and "beauty rating of professor from upper level male" for our model.

### Task 5: Multiple linear regression: one quantitative predictor, one qualitative predictor

```{r}
# Fitting the model
m_bty_gen <- lm(score ~ bty_avg + gender, data = evals)
tidy(m_bty_gen)

plot(m_bty_gen)
summary(m_bty_gen)
```

### Diagnostic plot:

From the Residuals vs Fitted plot, I do not see any distinctive pattern which says that this plot doesn’t have non-linear relationships. This suggests that we can assume linear relationship between the predictors and the outcome variables.

From the QQ-plot, we can say that most residual points are away from the straight dashed line. So the model is not normal.

From the Scale-Location plot, the residuals appear randomly spread around the horizontal line.

From the Residuals vs Leverage plot, there is no influential point near Cook's distance line.

Yes it is. Gender variable made beauty average even more significant as the computed p-valueis even smaller.

### Simplified equation of the line corresponding to female professors: (Female = 0)

Estimated score = beta0hat + beta1hat(bty_avg) + beta2hat * 0
                = 3.74733824 + 0.07415537(bty_avg)
 
Simplified equation of the line corresponding to male professors: (Male = 1)

Estimated score = beta0hat + beta1hat(bty_avg) + beta2hat * 1
                = 3.74733824 + 0.07415537(bty_avg) + 0.17238955

### Interpretation:

Intercept 3.747 is significantly different from zero.

First slope 0.074 is significantly different from 0. For male and female professors, a 1-unit increase in beauty score is associated with 0.074 increase in the average evaluation score.

Second slope 0.172 is significantly different from 0. For a male and a female professors with the same beauty scores, male professors’ average evaluation score is expected to be higher by 0.172 points.

For two professors with the same beauty rating, the male professor tends to have a higher course evaluation score.

```{r}
# Fitting the model with gender removed and rank added in
m_bty_rank <- lm(score ~ bty_avg + rank, data = evals)
tidy(m_bty_rank)

plot(m_bty_rank)
summary(m_bty_rank)
```

E(Y) = β0 + β1X1 + β2I2 + β3I3

Here Y = Average professor evaluation score,
X1 = Quantitative predictor variable,
X2 = Categorical predictor variable (which will further divide into I1, I2, I3)

I1 = Indicator variable for Teaching which is 0 (slope = 0),
I2 = Indicator variable for Tenure track, 
I3 = Indicator variable for Tenured

R simply codes the category that comes first alphabetically as a 0.

### Simplified equation of the line corresponding for teaching professors: (Teaching = 0, Tenure track = 0, Tenured = 0)

Estimated score = β0hat + β1hat(bty_avg) + β2hat * 0 +  β3hat * 0
                = 3.98155 + 0.06783(bty_avg) + 0 + 0

### Simplified equation of the line corresponding for Tenure track professors: (Tenure track = 1, Tenured = 0)

Estimated score = β0hat + β1hat(bty_avg) + β2hat * 1 +  β3hat * 0
                = 3.74733824 + 0.07415537(bty_avg) + (-0.16070) * 1 + 0
 
### Simplified equation of the line corresponding for Tenured professors: (Tenure track = 0, Tenured = 1)

Estimated score = β0hat + β1hat(bty_avg) + β2hat * 0 +  β3hat * 1
                = 3.74733824 + 0.07415537(bty_avg) + 0 + (-0.12623) * 1
                
### bty_avg x gender interaction

```{r}
m_int <- lm(score ~ bty_avg * gender, data = evals)
tidy(m_int)
```

Since the p-value associated with the interaction term is < 0.05. So there is a statistically significant interaction between beauty average rating and gender. In other words, there is evidence that a synergy effect exists between these 2 variables. Their combination is more powerful than the sum of their effects.

To study the effect of beauty average rating and gender on the professor evaluation score.

Y = β0 + β1 X1 + β2 X2 + β3 X1X2

Score = 3.95005984 + 0.03064259 * bty_avg - 0.18350903 * gender + 0.07961855 * bty_avg * gender

For Female: 

Score = 3.95005984 + 0.03064259 * bty_avg - 0.18350903 * 0 + 0.07961855 * bty_avg * 0
      = 3.950 + 0.0306 * bty_avg

For Male: 

Score = 3.95005984 + 0.03064259 * bty_avg - 0.18350903 * 1 + 0.07961855 * bty_avg * 1
      = 3.766 + 0.1102 * bty_avg
      
From last 2 equations, we can easily see that a 0.0306 unit increase in beauty average rating has a larger effect on prof evaluation score for those who are female (coefficient = 0.0306) compared to those who are male (coefficient = 0.1102).

